---
title: "Reproducible Research: Peer Assessment 1"
author: "Andrew Strasman"
output: 
html_document:
keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}
# activity.zip is already included in our clone of the repository 
unzip("activity.zip", overwrite = TRUE)
ds <- read.csv("activity.csv")

# coerce to factor
ds$interval <- as.factor(ds$interval)

# going to need these later for charting purposes
intervals <- as.numeric(levels(ds$interval))

```


## What is mean total number of steps taken per day?

```{r, echo=TRUE}
TotSteps <- tapply(ds$steps, ds$date, sum, na.rm=TRUE)

library(ggplot2)
qplot(TotSteps)

mean(TotSteps)
median(TotSteps)
```


## What is the average daily activity pattern?

```{r, echo=TRUE}

AvgSteps <- tapply(ds$steps, ds$interval, mean, na.rm=TRUE)
ds2 <- data.frame(time=names(AvgSteps),avgsteps=AvgSteps)

p <- ggplot(ds2, aes(intervals, avgsteps))
p <- p + geom_line(color="blue",lwd=2)
p <- p + ggtitle("Time series plot of the average number of steps taken")

p

```

```{r, echo=TRUE}
# Which time slice has the most average steps?
MaxSteps <- which.max(ds2$avgsteps)
names(MaxSteps)
```

## Imputing missing values

```{r, echo=TRUE}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
TotNa <- sum(is.na(ds$steps))
print(TotNa)
```


Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r, echo=TRUE}
# clone the original
ds3 <- ds

# fill in the NAs with means for that period
ds3$steps[is.na(ds3$steps)] <- tapply(ds3$steps, ds3$interval, mean, na.rm = TRUE)

```


```{r, echo=TRUE}
TotSteps2 <- tapply(ds3$steps, ds3$date, sum, na.rm=TRUE)

qplot(TotSteps2)

mean(TotSteps2)
median(TotSteps2)
```

Measure | Original | Modified | Difference
--------|----------|----------|------------
Mean    |   9354  |   10766  |   +1412  
Median  |   10395  |   10766  |    +371

### The impact of imputing missing data (with the mean of the prevailing time slice) has:
1. Raised both the mean and the median
2. Caused the mean and median to converge


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
# clone the original
ds3$day <- "weekday"
ds3$day[weekdays(as.POSIXlt(ds3$date)) == "Saturday"] <- "weekend"
ds3$day[weekdays(as.POSIXlt(ds3$date)) == "Saturday"] <- "weekend"
as.factor(ds3$day)

```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:

```{r, echo=TRUE}
#AvgSteps2 <- with(ds3[ds3$day == "weekend", ], tapply(ds3$steps, ds3$interval, mean))
#
#ds4 <- data.frame(time=names(AvgSteps2),avgsteps=AvgSteps2)
#AvgSteps3 <- with(ds3[ds3$day == "weekend", ], tapply(ds3$steps, ds3$interval, mean))


#p <- ggplot(ds4, aes(intervals, avgsteps))
#p <- p + geom_line(color="blue",lwd=2)
#p <- p + facet_grid(. ~ day)

## baby is sick and up crying with a double-ear infection,,,,  gotta call this a night...
library(lattice)
attach(ds3)
densityplot(~steps|day)

```

